### 2021-02-16 rscripts/manuscript-results.Rmd
update results with revision

---
title: "Different populations agree on which moral arguments underlie which opinions"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 7.1, fig.height = 3.15,
                      dpi = 300, dev="jpeg", tidy = TRUE)

library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(forcats)
library(ggplot2)
library(DescTools)
library(kableExtra)
library(cowplot)

theme_set(theme_half_open(11))

```

```{r data}

# US

us_arg <- read_csv("../data/us-argument-data-clean.csv")

us_arg <- us_arg %>% 
  mutate(
    sex = fct_relevel(sex, "male"),
    age_gr = cut(age, 
                 breaks = c(0, median(age, na.rm = TRUE), 100), 
                 labels = c("Younger", "Older")),
    verb_ability = ifelse(wordsum > median(wordsum, na.rm = TRUE),
                          "Higher",
                          "Lower"),
    verb_ability = fct_relevel(verb_ability, "Lower"),
    edu = fct_collapse(edu, 
                       Lower = "<bach", 
                       Higher = c("bach", "grad")),
   position = paste0(issue, "_", type),
   mf = factor(mf, levels = c("Harm", "Fairness", "Liberty", "Violence",
                            "Authority", "Purity", "Ingroup", "Government"))
  )

# UK

uk_arg <- read_csv("../data/uk-argument-data-clean.csv")

uk_arg <- uk_arg %>% 
  mutate(
    age_gr = cut(age, 
                 breaks = c(0, median(age, na.rm = TRUE), 100), 
                 labels = c("Younger", "Older")),
    polviews = cut(polviews_10, 
                   breaks = c(0, 3, 6, 10),
                   include.lowest = TRUE,
                   labels = c("Left", "Moderate", "Right")),
    edu = factor(edu, levels = c("lower", "higher"), labels = c("Lower", "Higher")),
    position = paste0(issue, "_", type),
   mf = factor(mf, levels = c("Harm", "Fairness", "Liberty", "Violence",
                            "Authority", "Purity", "Ingroup", "Government"))
  )

# Israel
il_arg <- read_csv("../data/il-argument-data-clean.csv")

il_arg <- il_arg %>% 
  mutate(
    age_gr = cut(age, 
                 breaks = c(0, median(age, na.rm = TRUE), 100), 
                 labels = c("Younger", "Older")),
    polviews = cut(polviews_10, 
                   breaks = c(0, 3, 6, 10),
                   include.lowest = TRUE,
                   labels = c("Left", "Moderate", "Right")),
    edu = ifelse(edu > 4, "Higher", "Lower"),
    position = paste0(issue, "_", type) 
  )

# Brazil
br_arg <- read_csv("../data/br-argument-data-clean.csv")
br_arg <- br_arg %>% 
  mutate(
    age_gr = cut(age, 
                 breaks = c(0, median(age, na.rm = TRUE), 100), 
                 labels = c("Younger", "Older")),
    polviews = cut(polviews_10, 
                   breaks = c(0, 3, 6, 10),
                   include.lowest = TRUE,
                   labels = c("Left", "Moderate", "Right")),
    edu = ifelse(edu > 4, "Higher", "Lower"),
    position = paste0(issue, "_", type)
  )

# Combined dataset
common_items <- read_csv("../data/gss-and-bsa-common-items.csv")

comb_arg <- bind_rows(
  us_arg %>% 
    filter(issue %in% common_items$gss_code) %>% 
    mutate(sample = "US"), 
  uk_arg %>% 
    filter(issue %in% common_items$bsa_code) %>% 
    # use GSS variable code for issue to unify between US and UK samples
    rename(bsa_code = issue) %>% 
    left_join(common_items %>% 
                select(issue = gss_code, bsa_code, reverse)) %>% 
    # one BSA item is reversed
    mutate(
      type = case_when(
        reverse == 1 & type == "pro" ~ "against", 
        reverse == 1 & type == "against" ~ "pro",
        TRUE ~ type
      ), 
      position = paste0(issue, "_", type),
      sample = "UK") %>% 
    select(-bsa_code, -reverse), 
  il_arg %>% 
    mutate(sample = "IL"), 
  br_arg %>% 
    mutate(sample = "BR")
)

```

```{r functions}

# estimate argument applicability within separate groups defined by group_var

link_by_group <- function(data, group_var) {

group_var <- enquo(group_var)

arg_link_by_group <- data %>% 
    drop_na(!!group_var) %>% 
    mutate(group = fct_relabel(!!group_var,
                               ~as.character(as.numeric(as.factor(.))))) %>%
    group_by(mf, position, group) %>% 
    summarise(link = mean(value)) %>% 
    spread(group, link)

arg_link_by_group %>% 
  drop_na() %>% 
  mutate(group = as_label(group_var))
}

# estimate CCC between argument applicability measures in the groups defined by group_var

ccc_by_group <- function(data, group_var) {
  group_var <- enquo(group_var)

  arg_link_by_group <- data %>% 
    drop_na(!!group_var) %>% 
    mutate(group = fct_relabel(!!group_var, ~as.character(as.numeric(as.factor(.))))) %>% 
    group_by(mf, position, group) %>% 
    summarise(link = mean(value), .groups = "drop") %>% 
    spread(group, link)

  arg_link_by_group %>% 
    drop_na() %>% 
    summarise(CCC(`2`, `1`)$rho.c,
              .groups = "drop") %>% 
    mutate(group = as_label(group_var))
}

get_group_prop <- function(data, group_var) {
  group_var <- enquo(group_var)

  data %>% 
    drop_na(!!group_var) %>% 
    summarise(prop = mean(!!group_var == first(!!group_var))) %>% 
    mutate(group = as_label(group_var))
}

estimate_random_group_ccc <- function(data, p = .5) {

  data %>% 
    # draw random group
    mutate(rand_gr = rbinom(n = n(), size = 1, prob = p),
           rand_gr = factor(rand_gr)) %>% 
    # aggregate arg connection by random group
    group_by(mf, position, rand_gr) %>%
    summarise(link = mean(value), .groups = "drop") %>%
    # calculate CCC
    spread(rand_gr, link) %>% 
    summarise(ccc = DescTools::CCC(`0`, `1`)$rho.c$est) %>% 
    pull(ccc)
}

```

# Study 1. US

## Method

```{r us_sample}

us_sample <- us_arg %>% 
  select(id, age, sex, polviews, edu, verb_ability, issue) %>% 
  distinct() %>% 
  group_by(id, age, sex, polviews, edu, verb_ability) %>% 
  summarise(n_issues = n()) %>% 
  ungroup()

us_sample %>% 
  summarise(n = n(),
            mean(n_issues),
            women = mean(sex == "female", na.rm = TRUE),
            mean_age = mean(age, na.rm = TRUE),
            sd_age = sd(age, na.rm = TRUE),
            liberal = mean(polviews == "liberal", na.rm = TRUE),
            conservative = mean(polviews == "conservative", na.rm = TRUE), 
            verb_higher = mean(verb_ability == "Higher"), 
            edu_higher = mean(edu == "Higher"))

```

## Results

```{r us_boxpl, fig.cap = "Figure 1. Boxplots showing how applicability, estimated in the entire sample of Study 1, of eight different kinds of moral arguments varied across 196 moral opinions. The box represents the interquartile (IQ) range with the dark line indicating the median. The whiskers reach the min and max values in case these are at most 1.5 times the box height outside the IQ range. Circles and stars signify outliers (values between 1.5 and 3 times the IQ range) and extreme outliers (more than 3 times the IQ range), respectively."}

us_arg_connection  <- us_arg %>% 
  group_by(mf, position) %>% 
  summarise(link = mean(value))

us_arg_connection <- us_arg_connection %>% 
  mutate(iqr = IQR(link), 
         lower = quantile(link, .25), 
         higher = quantile(link, .75),
         outl_type = case_when(
          link < lower - 3*iqr ~ 8, 
          link > higher + 3*iqr ~ 8,
          link < lower - 1.5*iqr ~ 1, 
          link > higher + 1.5*iqr ~ 1
          ))

us_arg_connection %>% 
  ggplot(aes(mf, link)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(shape = outl_type), size = 1.5, stroke = .5) +
  scale_shape_identity() +
  theme(legend.position = "none") +
  labs(y = "Arguments applicablity", x = "Moral foundation")

# ggsave("fig1_us-arg-applicability.jpeg", width = 180, height = 80, units = "mm", dpi = 300)

```

```{r us_ccc, cache = TRUE}

us_mf_agreement <- bind_rows(
  ccc_by_group(us_arg, polviews),
  ccc_by_group(us_arg, sex),
  ccc_by_group(us_arg, age_gr),
  ccc_by_group(us_arg, verb_ability),
  ccc_by_group(us_arg, edu)
)

us_group_p <- bind_rows(
  get_group_prop(us_arg, polviews),
  get_group_prop(us_arg, sex),
  get_group_prop(us_arg, age_gr),
  get_group_prop(us_arg, verb_ability),
  get_group_prop(us_arg, edu)
)

set.seed(28943)

us_mf_agreement <- left_join(us_mf_agreement, us_group_p)

make_1000_draws <- function(prop) {
  map_dbl(1:1000, ~estimate_random_group_ccc(us_arg, prop))
}

us_mf_agreement <- us_mf_agreement %>% 
  mutate(rand_ccc = map(prop, make_1000_draws))

us_mf_agreement <- us_mf_agreement %>% 
  mutate(p = map2_dbl(est, rand_ccc, ~mean(.y < .x)),
         mean_rand = map_dbl(rand_ccc, mean))

us_mf_agr_table <- us_mf_agreement %>% 
  mutate(label = sprintf("%.2f [%.2f, %.2f]", est, lwr.ci, upr.ci),
         p = sprintf("%.3f", p), 
         group = factor(group,
                        levels = c("sex", 
                                   "age_gr",
                                   "polviews", 
                                   "edu",
                                   "verb_ability"), 
                        labels = c("Sex", 
                                   "Age", 
                                   "Ideology", 
                                   "Education",
                                   "Verbal ability"))) %>% 
  select(group, label, mean_rand, p) %>% 
  arrange(group)

us_mf_agr_table %>%  
  kbl(col.names = c("Compared subsamples", 
                    "CCC with 95% CI", 
                    "Expected CCC based on random split", 
                    "p"),
      caption = "Table 2. CCC values with 95% confidence interval.",
      digits = 3) %>% 
  kable_classic(full_width = FALSE)

```

```{r fig.height=3.94, fig.cap="Figure 2. Scatter plots of 1,568 argument applicability scores (8 types of arguments by 196 moral opinions), measured in different groups in the United States: women vs. men (A),  younger vs. older (B), liberals vs. conservatives (C), higher vs. lower education (D), and higher vs. lower verbal ability (E). Regression lines in blue and reference lines for perfect agreement in black."}

us_mf_link_by_gr <- bind_rows(
  link_by_group(us_arg, polviews),
  link_by_group(us_arg, sex),
  link_by_group(us_arg, age_gr),
  link_by_group(us_arg, verb_ability),
  link_by_group(us_arg, edu)
)

scatter_plot <- function(data) {

  data %>% 
    ggplot(aes(`1`, `2`)) +
    geom_point(alpha = .2) +
    geom_abline() +
    geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(limits = c(0, 1), 
                       breaks = c(0, .25, .50, .75, 1))+ 
    scale_y_continuous(limits = c(0, 1), 
                       breaks = c(0, .25, .5, .75, 1)) +
    theme(axis.text = element_text(size = 9),
          plot.margin = margin(t = 1, r = 5, b = 1, l = 5))
}

pl1 <-  us_mf_link_by_gr %>% 
  filter(group == "sex") %>% 
  scatter_plot() +
  labs(x = "Women", y = "Men") +
  theme(strip.background = element_blank(), strip.text = element_blank())

pl2 <-  us_mf_link_by_gr %>% 
  filter(group == "age_gr") %>% 
  scatter_plot() +
  labs(x = "Younger", y = "Older")+
  theme(strip.background = element_blank(), strip.text = element_blank())

pl3 <- us_mf_link_by_gr %>% 
  filter(group == "polviews") %>% 
  scatter_plot() +
  labs(x = "Liberals", y = "Conservatives")

pl4 <- us_mf_link_by_gr %>% 
  filter(group == "edu") %>% 
  scatter_plot() +
  labs(x = "Lower education", y = "Higher education")+
  theme(strip.background = element_blank(), strip.text = element_blank())

pl5 <-  us_mf_link_by_gr %>% 
  filter(group == "verb_ability") %>% 
  scatter_plot() +
  labs(x = "Higher verbal ability", y = "Lower\nverbal ability")+
  theme(strip.background = element_blank(), strip.text = element_blank())

cowplot::plot_grid(pl1, pl2, pl3, pl4, pl5, 
                   labels = "AUTO", 
                   label_size = 11)

# ggsave("fig2_us-scatter-by-groups.jpeg", width = 180, height = 100,
#        units = "mm", dpi = 300)

```

Variance ratio between higher and lower verbal ability.

```{r}

us_mf_link_by_gr %>% 
  filter(group == "verb_ability") %>% 
  rename(higher = `1`, lower = `2`) %>% 
  group_by(mf) %>%
  summarise_at(vars(higher, lower), var) %>% 
  transmute(mf, ratio = lower/higher) %>% 
  arrange(ratio)

```

# Study 2. UK

## Method

```{r uk_sample}

uk_sample <- uk_arg %>% 
  select(polviews, id, issue, age, sex, edu) %>% 
  distinct() %>% 
  group_by(polviews, id, age, sex, edu) %>% 
  summarise(n_issues = n(), .groups = "drop")

uk_sample %>% 
  summarise(n = n_distinct(id),
            mean(n_issues),
            women = mean(sex == "Female", na.rm = TRUE),
            mean_age = mean(age, na.rm = TRUE),
            sd_age = sd(age, na.rm = TRUE),
            left = mean(polviews == "Left", na.rm = TRUE),
            moderate = mean(polviews == "Moderate", na.rm = TRUE),
            right = mean(polviews == "Right", na.rm = TRUE),
            edu_higher = mean(edu == "Higher", na.rm = TRUE))

```

## Results

```{r uk_boxpl, fig.cap = "Figure 3. Boxplots (defined in Figure 2) showing how applicability, estimated in the entire sample of Study 2, of eight different kinds of moral arguments varied across 216 moral opinions."}

uk_arg_connection  <- uk_arg %>% 
  group_by(mf, position) %>% 
  summarise(link = mean(value))

uk_arg_connection <- uk_arg_connection %>% 
  mutate(iqr = IQR(link), 
         lower = quantile(link, .25), 
         higher = quantile(link, .75),
         outl_type = case_when(
          link < lower - 3*iqr ~ 8, 
          link > higher + 3*iqr ~ 8,
          link < lower - 1.5*iqr ~ 1, 
          link > higher + 1.5*iqr ~ 1
          ))

uk_arg_connection %>% 
  ggplot(aes(mf, link)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(shape = outl_type), size = 1.5, stroke = .5) +
  scale_shape_identity() +
  theme(legend.position = "none") +
  labs(y = "Arguments applicablity", x = "Moral foundation")

# ggsave("fig3_uk-arg-applicability.jpeg", width = 180, height = 80, units = "mm", dpi = 300)

```

```{r uk_ccc, cache = TRUE}

uk_mf_agreement <- bind_rows(
  ccc_by_group(filter(uk_arg, 
                      polviews %in% c("Left", "Right")) %>% 
                 droplevels(), 
               polviews),
  ccc_by_group(uk_arg, sex),
  ccc_by_group(uk_arg, age_gr),
  ccc_by_group(uk_arg, edu)
)

uk_group_prop <- bind_rows(
  get_group_prop(filter(uk_arg, 
                      polviews %in% c("Left", "Right")) %>% 
                 droplevels(), polviews),
  get_group_prop(uk_arg, sex),
  get_group_prop(uk_arg, age_gr),
  get_group_prop(uk_arg, edu)
)

set.seed(653452)

uk_mf_agreement <- left_join(uk_mf_agreement, uk_group_prop)

make_1000_draws <- function(data, prop) {
  map_dbl(1:1000, ~estimate_random_group_ccc(data, prop))
}

uk_mf_agreement$data <- list(
   droplevels(filter(uk_arg, polviews %in% c("Left", "Right"))),
   uk_arg,
   uk_arg,
   uk_arg
)

uk_mf_agreement <- uk_mf_agreement %>% 
  mutate(rand_ccc = map2(data, prop, make_1000_draws))

uk_mf_agreement <- uk_mf_agreement %>% 
  mutate(p = map2_dbl(est, rand_ccc, ~mean(.y < .x)),
         mean_rand = map_dbl(rand_ccc, mean))

uk_mf_agr_table <- uk_mf_agreement %>% 
  mutate(label = sprintf("%.2f [%.2f, %.2f]", est, lwr.ci, upr.ci), 
         group = factor(group,
                        levels = c("sex", 
                                   "age_gr",
                                   "polviews", 
                                   "edu"), 
                        labels = c("Sex", 
                                   "Age", 
                                   "Ideology", 
                                   "Education")),
         p = sprintf("%.3f", p)) %>% 
  select(group, label, mean_rand, p) %>% 
  arrange(group)

uk_mf_agr_table %>%  
  kbl(col.names = c("Compared subsamples", 
                    "CCC with 95% CI", 
                    "Expected CCC based on random split", 
                    "p"),
      digits = 3,
      caption = "Table 3. The CCC value with 95% confidence interval.") %>% 
  kable_classic(full_width = FALSE)

```

```{r fig.width=3.35, fig.height=2.76, fig.cap="Figure 4. Scatter plots of 1,728 argument applicability scores (8 types of arguments by 216 moral opinions), measured in different groups in the United Kingdom: women vs. men (A),  younger vs. older (B), right-wing vs. left-wing (C), and higher vs. lower education (D). Regression lines in blue and reference lines for perfect agreement in black."}

uk_mf_diff <- bind_rows(
  link_by_group(filter(uk_arg, polviews %in% c("Left", "Right")) %>%
                  droplevels(), 
                polviews),
  link_by_group(uk_arg, sex),
  link_by_group(uk_arg, age_gr),
  link_by_group(uk_arg, edu)
)

scatter_plot <- function(data) {

  data %>% 
    ggplot(aes(`1`, `2`)) +
    geom_point(alpha = .2) +
    geom_abline() +
    geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(limits = c(0, 1), 
                       breaks = c(0, .25, .50, .75, 1)) + 
    scale_y_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1)) +
    theme_half_open(9) 
}

pl1 <-  uk_mf_diff %>% 
  filter(group == "sex") %>% 
  scatter_plot() +
  labs(x = "Women", y = "Men") +
  theme(strip.background = element_blank(), strip.text = element_blank())

pl2 <-  uk_mf_diff %>% 
  filter(group == "age_gr") %>% 
  scatter_plot() +
  labs(x = "Younger", y = "Older")+
  theme(strip.background = element_blank(), strip.text = element_blank())

pl3 <- uk_mf_diff %>% 
  filter(group == "polviews") %>% 
  scatter_plot() +
  labs(x = "Left-wing", y = "Right-wing")

pl4 <- uk_mf_diff %>% 
  filter(group == "edu") %>% 
  scatter_plot() +
  labs(x = "Lower education", y = "Higher education")+
  theme(axis.title.y = element_text(size = 8))

cowplot::plot_grid(pl1, pl2, pl3, pl4,
                   labels = "AUTO", 
                   label_size = 9)

# ggsave("fig4_uk-scatter-by-groups.jpeg", width = 85, height = 70, units = "mm", dpi = 300)

```

# Study 3. Combined US, UK, IL, and BR analysis

## Method

### IL sample

```{r il_sample}

il_sample <- il_arg %>% 
  select(id, issue, age, sex, polviews) %>% 
  distinct() %>% 
  group_by(id, age, sex, polviews) %>% 
  summarise(n_issues = n()) %>% 
  ungroup()

il_sample %>% 
  summarise(n = n(),
            mean(n_issues),
            women = mean(sex == "Female", na.rm = TRUE),
            mean_age = mean(age, na.rm = TRUE),
            sd_age = sd(age, na.rm = TRUE),
            left = mean(polviews == "Left", na.rm = TRUE),
            moderate = mean(polviews == "Moderate", na.rm = TRUE),
            right = mean(polviews == "Right", na.rm = TRUE))

```

### BR sample

```{r br_sample}

br_sample <- br_arg %>% 
  select(id, issue, age, sex, polviews) %>% 
  distinct() %>% 
  group_by(id, age, sex, polviews) %>% 
  summarise(n_issues = n()) %>% 
  ungroup()

br_sample %>% 
  summarise(n = n(),
            mean(n_issues),
            women = mean(sex == "Female", na.rm = TRUE),
            mean_age = mean(age, na.rm = TRUE),
            sd_age = sd(age, na.rm = TRUE),
            left = mean(polviews == "Left", na.rm = TRUE),
            moderate = mean(polviews == "Moderate", na.rm = TRUE),
            right = mean(polviews == "Right", na.rm = TRUE))

```

## Results

```{r ccc_bw_cntry, cache = TRUE}

comp_table <- tibble(combination = combn(unique(comb_arg$sample), 2, simplify = FALSE))

add_ccc_prop_for_comb <- function(combination) {
  df <- comb_arg %>% 
    filter(sample %in% combination)

  bind_cols(
    ccc_by_group(df, sample) %>% select(-group),
    get_group_prop(df, sample) %>% select(-group)
    )
}

comp_table <- comp_table %>% 
  mutate(label = map_chr(combination, paste0, collapse = " "),
         map_df(combination, add_ccc_prop_for_comb))

set.seed(83045)

make_1000_draws <- function(data, prop) {
  map_dbl(1:1000, ~estimate_random_group_ccc(data, prop))
}

comp_table <- comp_table %>% 
  mutate(rand_ccc = map2(combination, prop, 
                         ~make_1000_draws(filter(comb_arg, sample %in% .x), .y)))

comp_table <- comp_table %>% 
  mutate(p = map2_dbl(est, rand_ccc, ~mean(.y < .x)),
         mean_rand = map_dbl(rand_ccc, mean))

comp_table_format <- comp_table %>% 
  mutate(ccc = sprintf("%.2f [%.2f, %.2f]", est, lwr.ci, upr.ci),
         p = sprintf("%.3f", p)) %>% 
  select(label, ccc, mean_rand, p)

comp_table_format %>%  
  kbl(col.names = c("Compared subsamples", 
                    "CCC with 95% CI", 
                    "Expected CCC based on random split", 
                    "p"),
      digits = 3,
      caption = "Table 4. The CCC value with 95% confidence interval.") %>% 
  kable_classic(full_width = FALSE)

```

```{r com_scatter, fig.height = 4.7, fig.cap="Figure 5. Scatter plots of 432 argument applicability scores for (8 types of arguments by 54 moral opinions). Each row compares data from two of our four countries US, UK, Israel (IL) and Brazil (BR). Regression lines in blue and reference lines for perfect agreement in black."}

improve_plot <- function(pl, label_data) {

  pl +
    geom_point(alpha = .4) +
    geom_abline() +
    geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(limits = c(0, 1), 
                       breaks = c(0, .25, .50, .75, 1)) + 
    scale_y_continuous(limits = c(0, 1), 
                       breaks = c(0, .25, .5, .75, 1)) +
    theme(axis.text = element_text(size = 9),
          plot.margin = margin(t = 1, r = 5, b = 1, l = 5), 
          strip.text = element_text(size = 12))
}

arg_link_by_group <- comb_arg %>%
  group_by(mf, position, sample) %>%
  summarise(link = mean(value), .groups = "drop") %>%
  spread(sample, link)

us_uk_pl <- arg_link_by_group %>% 
  ggplot(aes(US, UK)) %>% 
  improve_plot(label_data = summarise(arg_link_by_group, 
                                      CCC(US, UK)$rho.c)) + 
  labs(x = "US sample", y = "UK sample")

us_il_pl <- arg_link_by_group %>% 
  ggplot(aes(US, IL)) %>% 
  improve_plot(label_data = summarise(arg_link_by_group, 
                                      CCC(US, IL)$rho.c)) + 
  labs(x = "US sample", y = "IL sample")

us_br_pl <- arg_link_by_group %>% 
  ggplot(aes(US, BR)) %>% 
  improve_plot(label_data = summarise(arg_link_by_group, 
                                      CCC(US, BR)$rho.c)) + 
  labs(x = "US sample", y = "BR sample")

uk_il_pl <- arg_link_by_group %>% 
  ggplot(aes(UK, IL)) %>% 
  improve_plot(label_data = summarise(arg_link_by_group, 
                                      CCC(UK, IL)$rho.c)) + 
  labs(x = "UK sample", y = "IL sample")

uk_br_pl <- arg_link_by_group %>% 
  ggplot(aes(UK, BR)) %>% 
  improve_plot(label_data = summarise(arg_link_by_group, 
                                      CCC(UK, BR)$rho.c)) + 
  labs(x = "UK sample", y = "BR sample")

il_br_pl <- arg_link_by_group %>% 
  ggplot(aes(IL, BR)) %>% 
  improve_plot(label_data = summarise(arg_link_by_group, 
                                      CCC(IL, BR)$rho.c)) + 
  labs(x = "IL sample", y = "BR sample")

cowplot::plot_grid(us_uk_pl, NULL, NULL,
                   us_il_pl, uk_il_pl, NULL,
                   us_br_pl, uk_br_pl, il_br_pl)

# ggsave("fig5_scatter-by-cntry.jpeg", width = 180, height = 120,
#        units = "mm", dpi = 300)

```

## Agreement on argument applicability

```{r agr_ccc, cache = TRUE}

comb_arg_full <- bind_rows(
  comb_arg %>% 
    filter(sample %in% c("IL", "BR")),
  us_arg %>%  mutate(sample = "US"),
  uk_arg %>% mutate(sample = "UK") 
)

# Probability to agree for each issue
p_agree <- comb_arg_full %>%
    filter(type == "pro") %>%
    select(sample, id, issue, opinion) %>%
    distinct() %>%
    group_by(sample, issue) %>%
    mutate(p = mean(opinion)) %>% 
    select(-opinion)

estimate_random_ccc <- function(data) {

  # Draw random opinion for each issue and espimated p
  rand_opin_simulated <- data %>% 
    mutate(pro = rbinom(n = n(), size = 1, prob = p),
           against = 1 - pro) %>%
    gather("type", "rand_opinion", pro, against)

  # Join argument data and calculate CCC
  left_join(rand_opin_simulated, 
            comb_arg_full %>% 
              select(sample, id, issue, mf, value),
            by = c("sample", "id", "issue")) %>% 
    group_by(sample, mf, issue, type, rand_opinion) %>%
    summarise(link = mean(value), .groups = "drop") %>%
    spread(rand_opinion, link) %>%
    drop_na() %>%
    group_by(sample) %>%
    summarise(est = CCC(`0`, `1`)$rho.c$est, .groups = "drop") 
}

set.seed(28943)

# Make 100 ccc simulations and average 
random_ccc <- map_df(1:1000, ~estimate_random_ccc(p_agree))

random_ccc_by_sample <- random_ccc %>% 
  chop(est) %>% 
  rename(rand_ccc = est)

ccc_for_agreement <- comb_arg_full %>% 
  group_by(sample, mf, issue, position, opinion) %>% 
  summarise(link = mean(value), .groups = "drop") %>% 
  spread(opinion, link) %>% 
  drop_na() %>% 
  group_by(sample) %>% 
  summarise(n_opinin = n_distinct(position), CCC(`0`, `1`)$rho.c)

ccc_for_agreement <- left_join(ccc_for_agreement, random_ccc_by_sample)

ccc_for_agreement <- ccc_for_agreement %>% 
  mutate(p = map2_dbl(est, rand_ccc, ~mean(.y < .x)),
         mean_rand = map_dbl(rand_ccc, mean),
         median_rand = map_dbl(rand_ccc, median))

ccc_for_agr_tbl <- ccc_for_agreement %>% 
  mutate(ccc = sprintf("%.2f [%.2f, %.2f]", est, lwr.ci, upr.ci), 
         sample = fct_rev(sample),
         mean_rand = sprintf("%.3f", mean_rand)) %>% 
  select(sample, n_opinin, ccc, mean_rand, p)

ccc_for_agr_tbl %>% 
  arrange(sample) %>% 
  kbl(col.names = c(
    "Country", 
    "Number of opinions",
    "CCC with 95% CI",
    "Expected CCC under perfect agreement",
    "p"
    ),
    caption = "Table 5. CCC values, with 95% confidence intervals, measuring in four countries the extent to which people with different opinions on the underlying moral issues agreed on argument applicability.") %>% 
  kable_classic(full_width = FALSE)

```